---
title: Public Indebtedness and Foreign Credit Demand for Emerging Bonds
author: "Augusto Netto, Gabriella Garcia and Maria Clara Drzeviechi"
output:
  html_document:
    df_print: paged
    highlight: haddock
    number_section: yes
    theme: readable
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    latex_engine: xelatex
    toc: yes
  word_document:
    toc: yes
---

<style>
body {
text-align: justify}
</style>


**Insper Data**



#Objective 

This study aims to investigate how a country’s level of public debt can be an obstacle to attract new foreign credit. If the country incurs high levels of debt, it is more likely that it will not be able to repay its creditor. Thus, in situations that a country’s level of public debt is too large, the creditors may choose to stop financing the country (Krugman, 1988), because of fearing a default. The objective is to verify the different effects of the level of public debt on the participation of foreign investors in government bonds and how the effects vary over time.



#Introduction

In the past 50 years, emerging and developing economies experienced four big indebtedness waves, three of which ended up in financial crises. In the 1980s, the association of low real interest rates and growing debt market led certain economies to raise considerably their levels of indebtedness; the result was the well-known Latin America’s debt crisis. A decade later, the world would watch the Asian financial crisis, due to the liberalization of financial markets and capital flows, allowing these countries to acquire loans in foreign currencies. Finally, in 2007-09, both emerging and advanced economies faced major recessions as a result of the global financial crisis.
 
It is important to notice that although happening in different decades and locations, these three episodes share a common denominator: they all started in periods with low real interest rates and an escalating indebtedness. This scenario made the risk premium rise and subsequently there was a sudden stop of capital flows. 
 
In 2010, the fourth global wave of debt started, and it was the largest and fastest one. In accordance with the aforementioned ones, interest rates were low since the Global Financial Crisis and investors were seeking assets with greater profitability. It is also reasonable to consider that this wave has not yet met its end. By 2018, according to the IMF, global debt has reached 226% of GDP, representing an amount of 188 trillions of dollars.  Emerging and developing economies saw indebtedness grow 54 p.p. in 8 years, reaching 170% of GDP. Low income countries reached 67% of GDP in 2018 – that figure was 48% in 2010. A different situation is seen on advanced economies, once they have maintained the 265% ratio of debt to GDP on the same level since 2010.
 
Global debt waves have shown the need for emerging and developing markets to develop their sovereign bond markets to facilitate public debt financing and management, primarily through the issuance of domestic bonds in local currencies. This growth was reflected in the increase in domestic participation in government bonds, and in the increase in foreign interest in the debt of these governments, which has become increasingly relevant in the local bond market.
 
The raise in foreign participation in the sovereign debt market is also related to many susceptible advantages, such as a decrease in government borrowing costs, due to the higher demand for these bonds, presented in several studies (Andritzky, 2012; Arslanalp and Poghosyan, 2014; Jaramillo and Zhang, 2013; Warnock and Warnock, 2009) and the diversification of the investor base, reflecting different characteristics of investors in terms of risk tolerance and trading reasons, which may increase the liquidity of government debt bonds in the secondary market (World Bank and IMF , 2001).
 
However, foreign investors tend to be relatively more sensitive to risk, particularly foreign private investors, and can be an unstable source of demand in times of stress, because they have a broader pool of assets in which they can invest. As a result, they may be less inclined to maintain their participation during these episodes (Broner et al., 2013).
 
The constant rise in indebtedness, observed in the current wave of global debt and intensified by the current crisis of COVID-19, may culminate in a sudden increase in risk premiums, if investors consider debt to GDP levels to be unsustainable (Blanchard 2019; Henderson 2019; Rogoff 2019a, b). If this happens, both capital inflows and outflows decrease during economic crises, with this effect being stronger in global crises (Broner et al., 2013).
 
In this context, this study aims to analyze the different impacts of public debt on the participation of foreign creditors in government bonds, depending on different scenarios and macroeconomic fundamentals. Given the negative relationship between the level of public debt and foreign investors already explored in the literature, we will investigate how this effect varies over time and conditions. 

Therefore, a worldwide analysis will be conducted using the database provided by Arslanalp and Takahiro Tsuda, previously affiliated with the Monetary and Capital Markets Department of the International Monetary Fund. 



# Literature Review

Discussions on how indebtedness can affect the macroeconomic variables of a country are widely addressed in the literature. The main topics are related to foreign investors; growth; debt overhang and debt tolerance.

More initially, countries’ ability to attract non-local investors was studied by Burger and Warnok (2005). The study showed that countries with a history of low inflation tend to have a more developed bond market.  Subsequently, Brutti and Sauré (2012) studied the importance of a derivatives market for the quality of a bond market.

In this study, it is important to highlight the developed literature Arslanap and Tsuda (2014). The authors, in addition to having compiled the database used in this article, carried out several analyzes on foreign participation in the bond market.  It was noted that, as of 2010, there was a great flow of foreign capital to the emerging securities market. In addition, they concluded that countries with a lower debt-to-GDP rate are more attractive to foreign investors.

The relationship between debt and growth was first found to be non-linear by Reinhart, Rogoff, and Savastano (2003). Later, Reinhart and Rogoff (2010) analyze countries distinguishing developed and emerging markets and they found out that, for both groups, a 90% debt to GDP ratio can be detrimental for growth. On the other hand, Kumar and Woo (2010) produced evidence that the negative impact of debt on growth is higher for developing countries when compared with developed countries. 

Thinking about debt tolerance, Reinhart, Rogoff and Savastano (2003) introduce the concept of debt intolerance and analyze how emerging markets find it difficult to face high levels of debt, while advanced countries face this issue more easily.  Later, Catão and Kapur (2006) discussed the macroeconomic determinants of a country’s volatility and how it impacts the spread rate it has to pay.

The concept of debt overhang refers to a debt burden so large that the country can not take any additional debt to finance itself. Krugman (1988) discussed the tradeoff for creditors when facing a debt overhang: financing or forgiving. Deshpande (1995) discusses how a situation of debt overhang can discourage investment. Reinhart, Reinhart, and Rogoff (2012) punctuated the main episodes in history about debt overhang.



#Data

The dataset used in this study is composed by the database provided by Arslanalp and Takahiro Tsuda, previously affiliated with the Monetary and Capital Markets Department of the International Monetary Fund. Arslanalp and Takahiro Tsuda provide estimates of investor holdings of government debt of 24 advanced economies and 24 emerging market economies. 

The investor base is grouped under six classes: domestic central bank, domestic banks, domestic nonbanks, foreign official sector, foreign banks, and foreign nonbanks. Banks include depository institutions other than central banks, based on the definition used in the IMF’s International Financial Statistics (IFS). Nonbanks include other investors, including insurance companies, pension funds, and investment funds. Investment funds could be mutual funds, exchange-traded funds (ETFs), or sovereign wealth funds (SWFs). The foreign official sector for Emerging Markets includes official loans and foreign central bank holdings of EM government debt as reserve assets, and for Advanced Markets includes foreign central banks and other foreign official creditors. 

For this study, 45 of the 48 countries present at the base of Arslanalp and Tsuda are used. The data are annualized through the fourth quarter of each year and the sample period used covers the years 2004 to 2019.

Macroeconomic and financial data were obtained from the World Economic Outlook Database published by the International Monetary Fund in April 2020. For countries where the nominal interest rate was incomplete on the basis of the IMF, the missing data was obtained on the website of the central banks in each country. 


# Methodology

The methodology used in this study will consist of regressions with panel data. Within and pooled models will be used and compared in order to obtain the most appropriate model. The advantages of using panel data are the possibility to follow the evolution of several countries over time, being able to capture the specific effects of each country and each period of time.

The regression response variable will be the proportion of foreign participation in the total debt issued by a given country. Regarding the explanatory variables, the main one will be the debt-to-GDP ratio. The other variables to control the regression are macroeconomic fundamentals, such as inflation, the balance of payments, exchange rate volatility, and others.



# Dataset Description

The dataset used in this study is composed by the database provided by Arslanalp and Takahiro Tsuda (2012 and 2014), which compiles comparable estimates of investor holdings of sovereign debt. In addition, data provided bt the World Economic Outlook from IMF and the WOrld Bank are used. The dataset is composed of 45 countries and covers sixteen years, from 2004 to 2019, quartely frequncy.  

```{r, echo=FALSE, message=FALSE} 

#adicionando libraries
rm(list=ls())
library(tidyverse)
library(scales)
library(ggthemes)
library(spData)
library(gganimate)
library(readr)
library(readxl)
library(spData)
library(plotly)
library(sf)
library(viridis)
library(ggcharts)
library(ggrepel)
library(cowplot)
library(gapminder)
library(plm)
library(stargazer)
library(lmtest)



# Puxando a dataset_total e a dataset_total_index
dataset_total <- read.csv("dataset_total.csv") 

dataset_total_index <- read.csv("dataset_total_index.csv") 


```



## Ranking 

```{r , echo=FALSE, message=FALSE}

  dataset_total %>%
  filter(year %in% c(2006, 2010, 2014, 2019), !is.na(develop)) %>%
  rename(Development = develop) %>% 
  bar_chart(x = country, y = debt_to_GDP, facet = year, top_n = 15, fill = Development) +
  labs( x = "", y = "Debt-to-GDP (%)",
  title = "Top 15 Debt-to-GDP Ratio", fill = "Development") +
  theme_classic() +
  scale_fill_manual("Development", values = c("EM" = "red4", "AM" = "navyblue"))+
  theme(legend.title = element_text(face = "bold", size = 10))



```



```{r , echo=FALSE, message=FALSE}

dataset_total %>%
  mutate(foreign_debt_perc=(foreign_debt/total_debt)*100) %>% 
    rename(Development = develop) %>% 
  filter(year %in% c(2004, 2009, 2014, 2019), !is.na(Development)) %>%
  bar_chart(x = country, y = (foreign_debt_perc), facet = year, top_n = 15, fill = Development) +  
  labs( x = "", y = "Foreign Investors (%)",
  title = "Top 15 Countries with Debt Held by Foreign Investors", fill = "Development") +
  theme_classic() +
  theme(legend.title = element_text(face = "bold", size = 10)) +
  scale_fill_manual("Development", values = c("EM" = "red4", "AM" = "navyblue"))


```



## Evolution of public debt in emerging markets

```{r , echo=FALSE, message=FALSE}

graph_debt_asia <- dataset_total %>% 
  filter(!is.na(develop)) %>% 
  rename(Countries = country) %>% 
  filter(develop=="EM", continent %in% c("Asia", "Africa")) %>% 
  ggplot(aes(x = year, y = debt_to_GDP, color=Countries)) +
  geom_smooth(se = FALSE)+
  labs(x = "Year", y = "Debt-to-GDP (%)", title = "Debt-to-GDP Ratio", subtitle = "Emerging Markets - Asia e Africa") +
     ylim(0,160)+
  theme_bw()


graph_debt_lamerica <- dataset_total %>% 
  filter(!is.na(develop)) %>% 
  rename(Countries = country) %>% 
  filter(develop=="EM", continent%in% c("South America", "North America")) %>% 
  ggplot(aes(x = year, y = debt_to_GDP, color=Countries)) +
  geom_smooth(se = FALSE)+
  labs(x = "Year", y = "Debt-to-GDP (%)", title = "Debt-to-GDP Ratio", subtitle = "Emerging Markets - Latin America") +
     ylim(0,100)+
  theme_bw()


graph_debt_europe <- dataset_total %>% 
  filter(!is.na(develop)) %>% 
    rename(Countries = country) %>% 
  filter(develop=="EM", continent=="Europe") %>% 
  ggplot(aes(x = year, y = debt_to_GDP, color=Countries)) +
  geom_smooth(se = FALSE)+
  labs(x = "Year", y = "Debt-to-GDP (%)", title = "Debt-to-GDP Ratio", subtitle = "Emerging Markets - Europe") +
   ylim(0,100)+
  theme_bw()


graph_debt_lamerica
graph_debt_europe
graph_debt_asia

```



## Evolution of the participation of foreign investors in government debt in Emerging Markets


```{r , echo=FALSE, message=FALSE}

graph_debt_foreign_asia <- dataset_total %>% 
  filter(!is.na(develop)) %>% 
  rename(Countries = country) %>% 
  filter(develop=="EM", continent %in% c("Asia", "Africa")) %>% 
  mutate(foreign_debt_perc=(foreign_debt/total_debt)*100) %>% 
  ggplot(aes(x = year, y = foreign_debt_perc, color=Countries)) +
  geom_smooth(se = FALSE)+
  labs(x = "Year", y = "Foreign Investors (%)", title = "Total Debt Held by Foreign Investors", subtitle = "Emerging Markets - Asia e Africa") +
    ylim(0,80)+
  theme_bw()


graph_debt_foreign_lamerica <- dataset_total %>% 
  filter(!is.na(develop)) %>% 
  rename(Countries = country) %>% 
  filter(develop=="EM", continent%in% c("South America", "North America")) %>% 
  mutate(foreign_debt_perc=(foreign_debt/total_debt)*100) %>% 
  ggplot(aes(x = year, y = foreign_debt_perc, color=Countries)) +
  geom_smooth(se = FALSE)+
  labs(x = "Year", y = "Foreign Investors (%)", title = "Total Debt Held by Foreign Investors", subtitle = "Emerging Markets - Latin America") +
    ylim(0,80)+
  theme_bw()


graph_debt_foreign_europe <- dataset_total %>% 
  filter(!is.na(develop)) %>% 
    rename(Countries = country) %>% 
  filter(develop=="EM", continent=="Europe") %>% 
  mutate(foreign_debt_perc=(foreign_debt/total_debt)*100) %>% 
  ggplot(aes(x = year, y = foreign_debt_perc, color=Countries)) +
  geom_smooth(se = FALSE)+
  labs(x = "Year", y = "Foreign Investors (%)", title = "Total Debt Held by Foreign Investors", subtitle = "Emerging Markets - Europe") +
      ylim(0,80)+
  theme_bw()


graph_debt_foreign_lamerica
graph_debt_foreign_europe
graph_debt_foreign_asia


```


# Stylized Facts

Firstly, some stylized facts will be explained in relation to our dataset.


## Countries have different indebtedness levels

```{r , echo=FALSE, message=FALSE}

mundo <- world %>%
  mutate(name_long = str_replace_all(string = name_long, pattern = "Russian Federation", replacement = "Russia"),
         name_long = str_replace_all(string = name_long, pattern = "Republic of Korea", replacement = "Korea"))

mundo <- mundo %>% 
  select(-c('continent':'gdpPercap')) %>% 
  rename(country = name_long) %>% 
  full_join(dataset_total, by = "country") %>% 
  mutate(debt_to_GDP = as.numeric(debt_to_GDP))

mundo %>% 
  ggplot() +
  geom_sf(data = mundo, aes(geometry = geom), fill = "lightgrey") +
  geom_sf(aes(fill = debt_to_GDP)) +
  scale_fill_viridis_c(name = "(%)", option = "magma", trans = "sqrt", direction = -1) +
  ggtitle("Debt to GDP") +
  labs(subtitle = "Mean of 2004 to 2019") +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.title = element_text(face = "bold", size = 10),
        legend.background = element_rect(fill = NA),
        plot.title = element_text(face = "bold", size = 17)) +
      coord_sf(ylim = c(-50, 90))

legend.title = element_text(colour="blue", size=10, 
                            face="bold")
  

```

As the map shown above makes clear, countries around the world have contrasting levels of debt to GDP. Nations in grey are not part of our sample. 



## Advanced markets are more indebted than emerging markets

```{r, echo=FALSE, message=FALSE}
 graph_debt_sep<-dataset_total %>% 
   filter(!is.na(develop)) %>% 
   rename(Development = develop) %>% 
  group_by(year, Development) %>% 
  summarise(debt_to_GDP_mean = mean(debt_to_GDP)) %>% 
ggplot(aes(x = year, y = debt_to_GDP_mean, color=Development )) +
   scale_color_manual(values = c("EM" = "red4", "AM" = "navyblue"))+
    geom_point() +
    geom_smooth()+
    labs(x = "Year", y = "Debt-to-GDP ratio (%)", title = "Average Debt-to-GDP Ratio", subtitle = "Advanced and Emerging Markets") +
      ylim(35,90)+
    theme_light()

graph_debt_sep

```


In emerging markets after a sharp drop in the level of debt to GDP until 2010, the average debt to GDP of emerging countries has been incrising. 

In contrast, in advanced economies debt to gdp grew rapidly after 2008, reaching the highest levels between 2014 and 2015. After 2015 the average debt to GDP started to decrease. 


## Foreign investors have a preference for allocating their resources in advanced markets

```{r, echo=FALSE, message=FALSE}
 
graph_debt_foreign_sep <- dataset_total %>% 
 filter(!is.na(develop)) %>% 
 rename(Development = develop) %>% 
mutate(foreign_debt_USD=(foreign_debt*fx)) %>% 
group_by(year, Development) %>% 
summarise(debt_foreign_mean = mean(foreign_debt_USD)) %>% 
ggplot(aes(x = year, y = debt_foreign_mean, color=Development )) +
 scale_color_manual(values = c("EM" = "red4", "AM" = "navyblue"))+
  geom_point() +
  geom_smooth()+
  labs(x = "Year", y = "Foreign Investors (Billion USD)", title = "Total Debt Held by Foreign Investors", subtitle = "Advanced and Emerging Markets") +
      ylim(0,650)+
  theme_bw()

graph_debt_foreign_sep
```


It is possible to observe that foreign investors have allocated much more resources in bonds from advanced markets. 

###
```{r, echo=FALSE, message=FALSE}
 
graph_debt_foreign_pp <- dataset_total %>% 
   filter(!is.na(develop)) %>% 
   rename(Development = develop) %>% 
  mutate(foreign_debt_perc=(foreign_debt/total_debt)*100) %>% 
  group_by(year, Development) %>% 
  summarise(debt_foreign_pp_mean = mean(foreign_debt_perc)) %>% 
ggplot(aes(x = year, y = debt_foreign_pp_mean, color=Development )) +
   scale_color_manual(values = c("EM" = "red4", "AM" = "navyblue"))+
    geom_point() +
    geom_smooth()+
    labs(x = "Year", y = "Foreign Investors (%)", title = "Total Debt Held by Foreign Investors", subtitle = "Advanced and Emerging Markets") +
      ylim(30,50)+
    theme_bw()

graph_debt_foreign_pp

```


The average foreign participation in the public debt of the advanced countries exceeded the average of the emerging countries in 2005.

Foreign participation in emerging markets suffered a sharp drop from the beginning of the period until mid-2010, when the decrease was reversed in a slight increase in foreign participation.

Whereas in advanced countries foreign participation was increasing until 2015 and after that period the trend turned to a fall.


## The participation of domestic investors in the composition of debt it is not invariant over time

```{r, echo=FALSE, message=FALSE}
 
graph_debt_domestic_sep <- dataset_total %>% 
   filter(!is.na(develop)) %>% 
   rename(Development = develop) %>% 
  mutate(domestic_debt_USD_prop=(domestic_debt/total_debt)*100) %>% 
  group_by(year, Development) %>% 
  summarise(debt_domestic_mean = mean(domestic_debt_USD_prop)) %>% 
ggplot(aes(x = year, y = debt_domestic_mean, color=Development )) +
   scale_color_manual(values = c("EM" = "red4", "AM" = "navyblue"))+
    geom_point() +
    geom_smooth()+
    labs(x = "Year", y = "Domestic Investors (%)", title = "Total Debt Held by Domestic Investors", subtitle = "Advanced and Emerging Markets" ) +
    ylim(50,70)+
    theme_bw()

 graph_debt_domestic_sep

```


In addition to the domestic participation in debt varying for both emerging and advanced countries, it is possible to infer that those variations are inverse.



```{r, echo=FALSE, message=FALSE, warning=FALSE}

graph_nominal_rate_sep <- dataset_total %>% 
   filter(!is.na(develop)) %>% 
   filter(!is.na(nominal_rate)) %>% 
   rename(Development = develop) %>% 
  group_by(year, Development) %>% 
  summarise(debt_nominal_mean = mean(nominal_rate)) %>% 
ggplot(aes(x = year, y = debt_nominal_mean, color=Development )) +
   scale_color_manual(values = c("EM" = "red4", "AM" = "navyblue"))+
    geom_point() +
    geom_line()+
    labs(x = "Year", y = "Nominal Rate (%)", title = "Nominal Rate (%) ", subtitle = "Advanced and Emerging Markets" ) +
    theme_bw()

graph_nominal_rate_sep





```


## There is no clear relationship between participation of foreign investor in government debt and inflation

```{r, echo=FALSE, message=FALSE, warning=FALSE}

graph_debt_inflation <-  dataset_total %>% 
  filter(!is.na(develop)) %>% 
  rename(Development = develop) %>% 
  ggplot() + 
  geom_point(aes(inflation_mean, (foreign_debt/total_debt), colour = Development)) +
  xlim(0,17)+
  ylim(0,1) +
  labs(x = "Annual Inflation Rate (%)", y = "Foreign Investors (%) ", title = "Government Debt Held by Foreign Investors X Inflation", subtitle = "45 Countries (2004 - 2019)" ) + 
  theme_bw()+
scale_colour_manual("Development", values = c("EM" = "red4", "AM" = "navyblue")) 

graph_debt_inflation

```


## There is no clear relationship between the participation of foreign investor in government debt and nominal interest rate

```{r, echo=FALSE, message=FALSE, warning=FALSE}

graph_debt_interest <-  dataset_total %>% 
  filter(!is.na(develop)) %>% 
  rename(Development = develop) %>% 
  ggplot() + 
  geom_point(aes(nominal_rate, (foreign_debt/total_debt), colour = Development)) +
  xlim(0,20)+
  ylim(0,1) +
  labs(x = "Nominal Interest Rate (%)", y = "Foreign Investors (%)", title = "Government Debt Held by Foreign Investors X Nominal Interest Rate", subtitle = "45 Countries (2004 - 2019)" ) + 
  theme_bw()+
  scale_colour_manual("Development", values = c("EM" = "red4", "AM" = "navyblue")) 

graph_debt_interest

```



#Econometrics

-Type 1 regression: total foreign debt as response variable for all countries in the database
-Type 2 regression: total foreign debt excluding official debt as response variable for all countries in the database
-Type 3 regression: total foreign debt as response variable for advanced markets
-Type 4 regression: total foreign debt excluding official debt as response variable for advanced markets
-Type 5 regression: total foreign debt as response variable for emerging markets
-Type 6 regression: total foreign debt excluding official debt as response variable for emerging markets

```{r, echo=FALSE, message=FALSE, warning=FALSE}




```

